Reporducible Data - Assignment #1
========================================================

## Project setup

### Load data
Set your working directory
```{r setwd, echo=TRUE}
setwd('C:/www/r/RepData_PeerAssessment1')
```


Load the "quantified self" activity data 
```{r load activity data}
data <- read.table( "activity.csv"
                    , sep = ","
                    , header = TRUE
                   )
    

```

Make sure you got the data:

```{r datacheck}
head(data)
```

### Load dependant libraries
Libraries include
* xtable
* lubridate

```{r load libraries}
library(xtable)
library(lubridate)
```


### Globally Preprocess Data
Add weekday to data
```{r add weekday}
data$dayofweek = wday(data$date)
```

Make sure you got the day of week was added to the sata:

```{r datacheck weekeday}
head(data)
```


## Questions and Answers

### Q1. What is mean total number of steps taken per day?
**Process the data**
Aggregate the of steps per day
```{r}
dailySteps <- aggregate(steps ~ date, data, FUN=sum)
```

**Explore the data**
```{r q1 explore, fig.height=4}
head(dailySteps)

hist( dailySteps$steps, breaks = 10, col="yellow" )
```

***Answer* - Mean steps taken per day:**
```{r q1 restult mean}
mean(dailySteps$steps, na.rm=T)
```

***Answer* - Median steps taken per day:**
```{r q1 restult median}
median(dailySteps$steps, na.rm=T)
```



### Q2. What is the average daily activity pattern?
**Process the data**
Aggregate the of steps per interval
```{r}
intervalSteps <- aggregate(steps ~ interval, data, FUN=mean)
```

**Explore the data**
```{r q1 explore q2a, fig.height=4}
head(intervalSteps)
```
**Q2aMake a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days**

```{r plot data, fig.height=4}
plot(intervalSteps$interval, intervalSteps$steps, type="l", col="orange", main = "Average number of steps taken")
```

**Q2b) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**

***Answer* - The Interval with the maximum number of steps**
```{r}
intervalSteps[ intervalSteps$steps == max(intervalSteps$steps), ]
```




### Q3) What is the impact of inputing missing data on the estimates of the total daily number of steps?



**Add missing data**  
Create a function to get the average steps per interval
```{r input missing function}
getIntervalAverageSteps <- function(interval) {

        avgSteps = intervalSteps$steps[ intervalSteps$interval == interval]
        avgSteps

}
```


**Input missing data**  
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r input missing data}
dataFilled <- data
dataFilled$steps[ is.na(dataFilled$steps) ] <- getIntervalAverageSteps( dataFilled$interval )
```

**Determine "Filled" total number of steps taken each day**
```{r}
dailyStepsFilled <- aggregate(steps ~ date, dataFilled, FUN=sum)
```


**Make a histogram of the total number of steps taken each day**
```{r q3 filled hist, fig.height=4}
hist( dailySteps$steps, breaks = 10, col="red" )
```

***Answer* - Mean steps taken per day:**
```{r q3 restult filled mean}
meanFilled <- mean(dailyStepsFilled$steps, na.rm=T)
```

```{r q3 restult filled mean out, echo=FALSE}
meanFilled
```

***Answer* - Median steps taken per day:**
```{r q1 restult filled median}
median(dailyStepsFilled$steps, na.rm=T)
```

**Food for tought:**  
Why does the *summary* function come out with differnet number
```{r q1 filled summary}
summary(dailyStepsFilled$steps, na.rm=T)
summary(dailyStepsFilled$steps, na.rm=F)
```

***Answer* - What is the impact of imputing missing data on the estimates of the total daily number of steps?**  
There was not a significant impact to adding the data.  The unfilled mean was `r meanFilled` the filled mean was `r meanFilled`.  The unfilled median was `r meanFilled` and the filled median was `r meanFilled`.


### Q4) Are there differences in activity patterns between weekdays and weekends?
**Create a new factor variable in the dataset with two levels -- "weekday" and "weekend"**  

Create a weekend dataset
```{r weekend dataset}
dataWeekend <- subset( dataFilled, dataFilled$dayofweek == 1 | dataFilled$dayofweek == 7 )
```
checkout our weekend data
```{r weekend check}
head(dataWeekend)
head( unique( wday(dataWeekend$dayofweek, label = T) ) )
```

Find the weekend average
```{r weekend average}
weekendIntervalSteps <- aggregate(steps ~ interval, dataWeekend, FUN=mean)
```

Create a weekday dataset
```{r weekday dataset}
dataWeekday <- subset( dataFilled, dataFilled$dayofweek != 1 & dataFilled$dayofweek != 7 )
```

checkout our weekday data
```{r weekday check}
head(dataWeekday)
head( unique( wday(dataWeekday$dayofweek, label = T) ) )
```

Find the weekday average
```{r weekday average}
weekdayIntervalSteps <- aggregate(steps ~ interval, dataWeekday, FUN=mean)
```

```{r plot compare plots}
par(mfrow = c(2, 1))
plot(weekendIntervalSteps$interval, 
     weekendIntervalSteps$steps, 
     type="l", 
     col="green",
     main="Weekend Steps")
plot(weekdayIntervalSteps$interval, 
     weekdayIntervalSteps$steps, 
     type="l", 
     col="blue",
     main="Weekeday Steps")
```